---
title: "Statistical Rethinking Exercise - Chapter 4"
output: pdf_document
author: Yue - Mar.4, 2018
fontsize: 11pt
header-includes:
- \usepackage{setspace}\onehalfspacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### install R package
#install.packages(c("coda", "mvtnorm","devtools"))
library(devtools)
#devtools::install_github("rmcelreath/rethinking")
library(rethinking)
library(MASS)
map <- rethinking::map
```

# chapter 4  
## Easy - model definition  
$y_i~Normal(\mu, \sigma)$ -- likelihood  
$\mu~Normal(0,10)$ -- $\mu$ prior   
$\sigma~Uniform(0,10)$ -- $\sigma$ prior   
2 parameters are in the posterior distribution  
Form of Bayes' theorem that includes the proper likelihood and priors. 

$y_i~Normal(\mu, \sigma)$ -- likelihood  
$\mu_i=\alpha+\beta x_i$ -- linear model  
$\alpha~Normal(0,10)$ -- $\alpha$ prior   
$\beta~Normal(0,1)$ -- $\beta$ prior   
$\sigma~Uniform(0,10)$ -- $\sigma$ prior   
3 parameters are in the posterior distribution  

## Medium  
4M1 Simulate observed heights from the prior  
$y_i~Normal(\mu, \sigma)$ -- likelihood  
$\mu~Normal(0,10)$ -- $\mu$ prior   
$\sigma~Uniform(0,10)$ -- $\sigma$ prior   

```{r}
  sample_mu <- rnorm( 1e4 , 0 , 10 )
  sample_sigma <- runif( 1e4 , 0 , 10 )
  sim_prior <- rnorm( 1e4 , sample_mu , sample_sigma )
  dens( sim_prior )
```  

4M2 Translate the model into map formula  
flist<- alist(
      y ~ dnorm( mu , sigma ) ,
      mu ~ dnorm( 0 , 10 ) ,
      sigma ~ dunif( 0 , 10 )
    )  

4M3 Translate the map formula into a mathematical model definition  
flist<- alist(
      y ~ dnorm( mu , sigma ) ,
      mu <- a + b*x, 
      a ~ dnorm( 0 , 50 ) ,
      b ~ dunif( 0 , 10 ) ,
      sigma ~ dunif( 0 , 50 )
    )  
$y_i~Normal(\mu_i, \sigma)$ -- likelihood  
$\mu_i=\alpha+\beta x_i$ -- linear model  
$\alpha~Normal(0,50)$ -- $\alpha$ prior   
$\beta~Uniform(0,10)$ -- $\beta$ prior   
$\sigma~Uniform(0,50)$ -- $\sigma$ prior   

4M4 height model  
height.model <- map(
        alist(
            height ~ dnorm( mu , sigma ) ,
            mu <- a + b*year, 
            a ~ dnorm( 115 , 8 ) , #6-yr olds, white boys
            b ~ dnorm( 5 , 2 ) , 
            sigma ~ dunif( 0 , 5 )
        ) ,
        data=sample )  

4M5 If average height in first year=120cm, student got taller each year. Does this information lead you to change your choice of priors? Why?  
set: a ~ dnorm( 120 , 8 )  

4M6 variance among heights for students of the same age is never more than 64 cm, how does this lead you to revise your priors?  
set: sigma ~ dunif( 0 , 64 ) ?  

## Hard  
4H1 predict height with 89% intervals for individuals given weight   

```{r}
# get height data
data(Howell1)
d <- Howell1

#plot raw data
plot(d$height~d$weight)

# build model using map
m1 <- map(
  alist(
    height ~ dnorm (mu, sigma),
    mu <- a + b * weight, 
    a ~ dnorm(165, 80),
    b ~ dnorm(0,10),
    sigma ~ dunif(0,64) #given information in 4M6
  ),
  data = d
)

# simulate height based on model 
sample.weight <-as.data.frame(cbind(c(1:5),c(46.95,43.72,64.78,32.59,54.63)))
colnames(sample.weight)<-c("id","weight")

# use link to compute mu for each sample from posterior and for each weight
mu <- link( m1 , data=data.frame(weight=sample.weight$weight))
# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
mu.PI <- apply( mu , 2 , PI , prob=0.89 )

# simulate height
sim.height <- sim( m1 , data=list(weight=sample.weight$weight) , n=1e4 )
# summarize the distribution of simulation
sim.mean <- apply( sim.height , 2 , mean )
sim.HPDI <- apply( sim.height , 2 , HPDI , prob=0.89 )
sim.PI <- apply( sim.height , 2 , PI , prob=0.89 )

cbind(sim.mean, mu.mean)
cbind(sim.HPDI, mu.HPDI, sim.PI, mu.PI) #almost the same

#plot adding MAP line and HPDI
plot(d$height[d$weight>=30]~d$weight[d$weight>=30])
lines( sample.weight$weight, sim.mean , col="red") # draw MAP line
shade( sim.HPDI , sample.weight$weight, col=col.alpha("red",0.5) ) # draw HPDI region for line
#shade( sim.PI , sample.weight$weight, col=col.alpha("green",0.3) ) # draw PI region for simulated heights

# check 1st individual
posterior <- extract.samples(m1)
sim1 <- rnorm(n = 1e4, mean = posterior$a + posterior$b*sample.weight$weight[1], sd = posterior$sigma)
mean(sim1)
PI(samples = sim1, prob = .89)
```  

4H2 below 18yr sample  
```{r}
# get below 18 year subset
d2 <- Howell1[Howell1$age<18,]
dim(d2)
plot(height~weight, data=d2)
```  

(a) Linear regression using map.  
```{r}
m2 <- map(
        alist(
            height ~ dnorm( mu , sigma ) ,
            mu <- a + b*weight, 
            a ~ dnorm( 110 , 80 ) ,
            b ~ dnorm( 0 , 10 ) ,
            sigma ~ dunif( 0 , 64 )
        ) ,
        data=d2 )
m2
precis(m2, prob=0.95)
post<-extract.samples(m2)
mean(post$b)*10 # 10 units of increase in weight
```  
(b) Plot the raw data height ~ weight (see above).  
Superimpose the MAP regression line and 89% HPDI for the mean.

```{r}
# simulate mu
  weight.seq <- seq(from = 1, to = 50, length.out = 100)
  
  #method 1 using link function
  mu <- link( m2 , data=data.frame(weight=weight.seq) )
  mu.mean <- apply( mu , 2 , mean )
  #method 2 define linear model
  mu.link <- function(weight) post$a + post$b*weight
  mu2 <- sapply( weight.seq , mu.link )
  mu2.mean <- apply( mu2 , 2 , mean ) #almost the same
  
  #calculate PI
  mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
  mu.PI <- apply( mu , 2 , PI , prob=0.89 )
  
# predict height
  sim.height <- sim( m2 , data=list(weight=weight.seq) , n=1e4 )
  # summarize the distribution of simulation
  sim.mean <- apply( sim.height , 2 , mean )
  sim.HPDI <- apply( sim.height , 2 , HPDI , prob=0.89 )
  sim.PI <- apply( sim.height , 2 , PI , prob=0.89 )
  
#plot adding MAP line and HPDI
  plot(height~weight, data=d2)
  lines( weight.seq, mu.mean , col="red") # draw MAP line
  shade( mu.HPDI , weight.seq, col=col.alpha("red",0.5) ) # draw HPDI region for mean
  shade( sim.HPDI , weight.seq, col=col.alpha("yellow",0.3) ) # draw HPDI region for predicted height

```  

```{r, include=FALSE}
# b) - approach 2
library(MASS)
trials <- 1e5

weight.seq <- seq(from = 1, to = 45, length.out = 50)

# simulate mu then compute mean and hpdi
posterior.samples <- data.frame( mvrnorm(n = trials, mu = coef(m2), Sigma = vcov(m2)) )
mu.link <- function(weight) posterior.samples$a + posterior.samples$b * weight
mu <- sapply(X = weight.seq, FUN = mu.link)
mu.mean <- apply(X = mu, MARGIN = 2, FUN = mean)
mu.hpdi <- apply(X = mu, MARGIN = 2, FUN = HPDI, prob = .89)

# simulate heights then compute hpdi
height.link <- function(weight) rnorm(n = nrow(posterior.samples), mean = mu.link(weight), sd = posterior.samples$sigma)
height.samples <- sapply(X = weight.seq, FUN = height.link)
height.hpdi <- apply(X = height.samples, MARGIN = 2, FUN = HPDI, prob = .89)

# plot results
plot(height ~ weight, data = d2, col = col.alpha(rangi2, .5))
lines(x = weight.seq, y = mu.mean)
shade(object = mu.hpdi, lim = weight.seq)
shade(object = height.hpdi, lim = weight.seq)
``` 

Linear model height~weight does not seem a good fit given the data and graph above. Try linear model with transformation.  

4H3 log weight as the predictor  

(a) fit model  
$h_i~Normal(\mu_i, \sigma)$ -- likelihood  
$\mu_i=\alpha+\beta log(w_i)$ -- linear model  
$\alpha~Normal(178,100)$ -- $\alpha$ prior   
$\beta~Normal(0,100)$ -- $\beta$ prior   
$\sigma~Uniform(0,50)$ -- $\sigma$ prior   

```{r}
m3 <- map(
        alist(
            height ~ dnorm( mu , sigma ) ,
            mu <- a + b*log(weight), 
            a ~ dnorm( 178 , 100 ) ,
            b ~ dnorm( 0 , 100 ) ,
            sigma ~ dunif( 0 , 50 )
        ) ,
        data=d )
m3
precis(m3, prob=0.95)
invisible(post3 <- extract.samples(m3))
```  
model: height = `r post3$a` + `r post3$b` * log(weight)  

(b)  plot: use samples from the quadratic approximate posterior of the model in (a)  

1. superimpose the predicted mean height as a function of weight  
2. the 97% HPDI for the mean  
3. the 97% HPDI for predicted height  
```{r}
# simulate mu
  weight.seq <- seq(from = 1, to = max(d$weight), length.out = ((max(d$weight)-1)*2) )

  mu3 <- sapply(weight.seq, function(weight) post3$a + post3$b * log(weight) ) #model
  mu3.mean <- apply( mu3 , 2 , mean )
  mu3.HPDI <- apply( mu3 , 2 , HPDI , prob=0.97 )   #97% HDPI of the mean
  
# predict height
  sim3 <- sim( m3 , data=list(weight=weight.seq) , n=1e4 )
  sim3.HPDI <- apply( sim3 , 2 , HPDI , prob=0.97 )   #97% HDPI of predicted height
  
#plot adding MAP line and HPDI
  plot( height~weight, data=d, col=col.alpha(rangi2, 0.4)) #R code 4.73
  lines( weight.seq, mu3.mean , col="red") # draw MAP line
  shade( mu3.HPDI , weight.seq, col=col.alpha("red",0.5) ) # draw HPDI region for mean
  shade( sim3.HPDI , weight.seq, col=col.alpha("yellow",0.3) ) # draw HPDI region for predicted height
```  
The height ~ log(weight) seems a much better fit comparing to the height~weight model.  