---
title: "Statistical Rethinking Exercise - Chapter 3"
output: pdf_document
author: Yue - Feb.10, 2018
fontsize: 11pt
header-includes:
- \usepackage{setspace}\onehalfspacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, include=FALSE}
### install R package
#install.packages(c("coda", "mvtnorm","devtools"))
library(devtools)
#devtools::install_github("rmcelreath/rethinking")
library(rethinking)
```

# chapter 3 - Global tossing problem
## Easy  
```{r}
## grid approximation - 
p_grid<- seq( from=0, to=1, length.out=1000) # define grid
prior <- rep(1, 1000) # define prior
w<-6
n<-9
likelihood <- dbinom(w, size=n, prob=p_grid) #compute likelihood at each value in grid
unstd.posterior <- likelihood * prior # compute product of lieklihood and prior
posterior <- unstd.posterior / sum(unstd.posterior) # standardize the posterior
plot(posterior ~ p_grid, type = "l",col="darkgreen")

set.seed(100)
samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE) # draw 10000 samples from vector (p_grid) with sample distribution = exact posterior density

#plot density estimate
par(mfrow=c(1,2))
plot(samples) # y=proportion, x=#sample
dens(samples) # y=density, x=proportion
```  
3E1-3. intervals of defined boundaries  
```{r}
# add up posterior probability given boundaries
sum(posterior[p_grid<0.2]) 
sum(samples<0.2) / 1e4
sum(samples<0.8) / 1e4
sum(samples>0.2 & samples<0.8) / 1e4
```  
3E4-5. intervals of defined mass  
```{r}
# boundary given posterior probability mass
quantile(samples, 0.2)
quantile(samples, (1-0.2))
quantile(samples, c(0.2,0.8)) #specify lower/upper bound
```
3E6-7. Percentile intervals  
```{r}
HPDI(samples, prob=0.66) #highest posterior density interval - narrowest
PI(samples, prob=0.66) # PI - central
```  

## Medium  
**Global tossing example**  
3M1 Grid approximation, with 8 water out of 15 tosses data   
```{r}
w<-8
n<-15
p_grid<- seq( from=0, to=1, length.out=1000) # define grid
prior <- rep(1, 1000) # define prior, flat
likelihood2 <- dbinom(w, size=n, prob=p_grid) #compute likelihood at each value in grid
unstd.posterior2 <- likelihood2 * prior # compute product of lieklihood and prior
posterior2 <- unstd.posterior2 / sum(unstd.posterior2) # standardize the posterior
```  
3M2 HPDI  
```{r}
set.seed(100)
samples2 <- sample(p_grid, prob=posterior2, size=1e4, replace=TRUE) # 10000 samples
dens(samples2) # y=density, x=proportion

HPDI(samples2, prob=0.90) #HPDI 90%
```  
3M3-4 Posterior predictive check:  
simulate the distribution of samples, averaging over the posterior uncertainty in p  
```{r}
#3M3
postpred<-rbinom(1e4, size=15, prob=samples2)
simplehist(postpred)
mean(postpred==8)

#3M4
mean(rbinom(1e4, size=9, prob=samples2)==6)
```  
3M5  
```{r}
p_grid<- seq( from=0, to=1, length.out=1000) # define grid
prior2 <- ifelse(test=p_grid<0.5, yes=0, no=1) # define prior, 0 if p<0.5, constant if p>=0.5
likelihood2 <- dbinom(w, size=n, prob=p_grid) #compute likelihood at each value in grid
unstd.posterior3 <- likelihood2 * prior2 # compute product of lieklihood and prior
posterior3 <- unstd.posterior3 / sum(unstd.posterior3) # standardize the posterior
set.seed(100)
samples3 <- sample(p_grid, prob=posterior3, size=1e4, replace=TRUE) # 10000 samples
par(mfrow=c(2,2))
plot(prior2~p_grid, ylim=c(0, 1.5))
plot(prior~p_grid, ylim=c(0, 1.5))
dens(samples3)
dens(samples2)

HPDI(samples3, prob=0.90) #HPDI 90%

#Predictive check
par(mfrow=c(1,1))
postpred2<-rbinom(1e4, size=15, prob=samples3)
simplehist(postpred2, col="Blue", xlab="Tosses")
par(new=TRUE)
simplehist(postpred, axes=FALSE, col="Red", xlab="")
legend("topright", pch="|", c("new","old"), col=c("Blue","Red"))
```  

## Hard  
Data source: https://github.com/rmcelreath/rethinking/tree/master/data  
1=boy, 0=girl;
```{r}
birth1 <- c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
1,0,1,1,1,0,1,1,1,1)
birth2 <- c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
0,0,0,1,1,1,0,0,0,0)

data(homeworkch3)
boy<-sum(birth1) + sum(birth2) #total boys
n<-length(birth1) + length(birth2)

child<-cbind(birth1, birth2)
```  

3H1 posterior distribution for the probability of a birth being a boy  
```{r}
p_grid <- seq( from=0, to=1, length.out=1000) # define grid
prior <- rep(1, 1000) # uniform prior
likelihood <- dbinom(boy, size=n, prob=p_grid) 
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior) # standardize the posterior
plot(posterior ~ p_grid, type = "l",col="darkgreen")

p_grid[which.max(posterior)] # parameter value max the posterior probability
```  

3H2  
```{r}
set.seed(100)
samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE) # 10000 samples
HPDI(samples, prob=c(0.50,0.89,0.97)) 
```  

3H3  
```{r}
postpred <- rbinom(1e4, size=200, prob=samples)
simplehist(postpred)
dens(postpred, col="darkgreen")
abline(v=boy,col="red")
```

3H4 first birth  
```{r}
boy1<-sum(birth1)
n1 <- length(birth1)
likelihood1 <- dbinom(boy1, size=n1, prob=p_grid)
unstd.posterior1 <- likelihood1 * prior
posterior1 <- unstd.posterior1 / sum(unstd.posterior1) 
p_grid[which.max(posterior1)]
set.seed(100)
samples1 <- sample(p_grid, prob=posterior1, size=1e4, replace=TRUE) # 10000 samples
postpred1 <- rbinom(n = boy1, size = n1, prob = samples1)
simplehist(postpred1)
par(mfrow=c(1,2))
dens(postpred1, col="darkgreen", xlab="first birth posterior")
abline(v=boy1,col="red")

#use the 3H3 sample
postpred2 <- rbinom(n = boy1, size = n1, prob = samples)
dens(postpred2, col="darkblue", xlab="all birth posterior")
abline(v=boy1,col="red")

```  
Does not seem to fit well.  

3H5  
```{r}
boy2<-sum(birth2[birth1 == 0])
n2<-length(birth2[birth1 == 0])

likelihood2 <- dbinom(boy2, size=n2, prob=p_grid)
unstd.posterior2 <- likelihood2 * prior
posterior2 <- unstd.posterior2 / sum(unstd.posterior2) 
p_grid[which.max(posterior2)]
set.seed(100)
samples2 <- sample(p_grid, prob=posterior2, size=1e4, replace=TRUE) # 10000 samples
postpred2 <- rbinom(n = boy2, size = n2, prob = samples2)
simplehist(postpred2)
dens(postpred2, col="darkgreen")
abline(v=boy2,col="red")

plot(posterior2 ~ p_grid, type = "l",col="darkgreen")
```  
