---
title: "Notes on the book Statistical Rethinking"
author: Jason
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
  - \usepackage{tikz}
  - \usepackage{ifthen}
output:
  pdf_document:
    highlight: tango
fontsize: 12pt
geometry: margin = 1in
linkcolor: red
urlcolor: blue
---

```{r, include = FALSE, echo = FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
read_chunk('jasonCh2.R')
```

\newcommand{\rc}{\tikz\draw[black,fill=red] (0,0) circle (.5ex); }

\newcommand{\bc}{\tikz\draw[black,fill=blue] (0,0) circle (.5ex); }


# **Chapter 1: The Golem of Prague**

# **Chapter 2: Small Worlds and Large Worlds**

**Useful Example.**  Suppose there is a bag containing marbles.  We *know* there are four marbles in the bag and that each marble
may be either \textcolor{red}{red} or \textcolor{blue}{blue}.  We get some data in the form of \bc \rc \bc
We can think of the likelihood as way to count all of the possible ways of producing this sample **given** a conjecture of what
the bag looks like (i.e., a paramter for our model):

| **Conjecture**   | **Ways of producing sample** |
|:-----------------|:----------------------------:|
| \bc \bc \bc \bc | $4 \times 0 \times 4 = 0$ |

```{r, gardenPlot, echo = FALSE, out.width = '100%', fig.height=4, fig.width=6}
d
```
