---
title: "Notes on the book Statistical Rethinking"
author: Jason
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
  - \usepackage{tikz}
  - \usepackage{ifthen}
output:
  pdf_document:
    highlight: tango
fontsize: 12pt
geometry: margin = 1in
linkcolor: red
urlcolor: blue
---

```{r, include = FALSE, echo = FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
read_chunk('jasonCh2.R')
```

\newcommand{\rc}{\tikz\draw[black,fill=red] (0,0) circle (.5ex); }

\newcommand{\bc}{\tikz\draw[black,fill=blue] (0,0) circle (.5ex); }


# **Chapter 1: The Golem of Prague**

* Limitations of *deductive falsification*
  1. "Hypotheses are not models. The relations among hypotheses and different kinds of models are complex. Many models correspond
  to the same hypothesis, and many hypotheses correspond to a single model. This makes strict falsification impossible." (p. 4)
  2. "Measurement matters. Even when we think the data falsify a model, another ob- server will debate our methods and measures.
  They don’t trust the data. Sometimes they are right." (p.4)
* Main topics covered in this book...
  1. Bayesian data analysis
  2. Multilevel models
  3. Model comparison using information criteria

# **Chapter 2: Small Worlds and Large Worlds**

**Useful Example.**  Suppose there is a bag containing marbles.  We *know* there are four marbles in the bag and that each marble
may be either \textcolor{red}{red} or \textcolor{blue}{blue}.  Next, suppose we observe some data in the form of \bc \rc \bc in
that exact order (*so sequence matters*). We can think of the likelihood as a way to count all of the possible ways of producing
this sample **given** a conjecture of what the bag looks like (i.e., a paramter for our model).

* Consider one possible conjecture: \bc \rc \rc \rc .  How likely is the observed sample \bc \rc \bc given the conjecture? To answer
this question, we count the possible ways our conjecture can generate the sample.  Then we will compare the total number of ways across
different conjectures as a way to evaluate which conjecture is the most likely candidate for generating \bc \rc \bc .  The following
plot shows all of the possible samples that (the conjecture) \bc \rc \rc \rc can produce:

```{r, gardenPlot, echo = FALSE, out.width = '100%', fig.height=4, fig.width=6}
d
```
From this figure we see that there are 3 possible ways to generate the observed data. The following table lists all of the ways that
the 5 possible conjectures could generate the sample \bc \rc \bc:

| **Conjecture**   | **# of ways to produce** \bc \rc \bc |
|:-----------------|:--------------------------------------:|
| \bc \rc \rc \rc | $1 \times 3 \times 1 = 3$ |
| \bc \bc \rc \rc | $2 \times 2 \times 2 = 8$ |
| \bc \bc \bc \rc | $3 \times 1 \times 3 = 9$ |
| \bc \bc \bc \bc | $4 \times 0 \times 4 = 0$ |
| \rc \rc \rc \rc | $0 \times 4 \times 0 = 0$ |

* Prior belief.

* Bayesian updating & posterior distribution.

# **Chapter 3**

# **Chapter 4**
