---
title: "Notes on the book Statistical Rethinking"
author: Jason
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{tikz}
  - \usepackage{ifthen}
output:
  pdf_document:
    fig_width: 5
    fig_height: 4
    fig_caption: true
    highlight: tango
fontsize: 12pt
geometry: margin = 1in
linkcolor: red
urlcolor: blue
---

```{r, include = FALSE, echo = FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
```

\newcommand{\rc}{\tikz\draw[black,fill=red] (0,0) circle (.5ex); }

\newcommand{\bc}{\tikz\draw[black,fill=blue] (0,0) circle (.5ex); }


# **Chapter 1: The Golem of Prague**

# **Chapter 2: Small Worlds and Large Worlds**

**Useful Example.**  Suppose there is a bag containing marbles.  We *know* there are four marbles in the bag and that each marble
may be either \textcolor{red}{red} or \textcolor{blue}{blue}.  We get some data in the form of \bc \rc \bc
We can think of the likelihood as way to count all of the possible ways of producing this sample **given** a conjecture of what
the bag looks like (i.e., a paramter for our model):

| **Conjecture**   | **Ways of producing sample** |
|:-----------------|:----------------------------:|
| \bc \bc \bc \bc | $4 \times 0 \times 4 = 0$ |

```{r, echo = FALSE}
d <-
         tibble(position = c((1:4^1) / 4^0,
                             (1:4^2) / 4^1,
                             (1:4^3) / 4^2),
                draw     = rep(1:3, times = c(4^1, 4^2, 4^3)),
                fill     = rep(c("b", "w"), times = c(1, 3)) %>%
                  rep(., times = c(4^0 + 4^1 + 4^2)))

lines_1 <-
                          tibble(x    = rep((1:4), each = 4),
                                 xend = ((1:4^2) / 4),
                                 y    = 1,
                                 yend = 2)

lines_2 <-
                                   tibble(x    = rep(((1:4^2) / 4), each = 4),
                                          xend = (1:4^3)/(4^2),
                                          y    = 2,
                                          yend = 3)

d %>%
                                            ggplot(aes(x = position, y = draw)) +
                                            geom_segment(data  = lines_1,
                                                         aes(x = x, xend = xend,
                                                             y = y, yend = yend),
                                                         size  = 1/3) +
                                            geom_segment(data  = lines_2,
                                                         aes(x = x, xend = xend,
                                                             y = y, yend = yend),
                                                         size  = 1/3) +
                                            geom_point(aes(fill = fill),
                                                       shape = 21, size = 3) +
                                            scale_y_continuous(breaks = 1:3) +
                                            scale_fill_manual(values  = c("navy", "white")) +
                                            theme(panel.grid.minor = element_blank(),
                                                  legend.position = "none")
```
