---
title: "Notes on the book Statistical Rethinking"
author: Jason
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
  - \usepackage{tikz}
  - \usepackage{ifthen}
output:
  pdf_document:
    highlight: tango
fontsize: 12pt
geometry: margin = 1in
linkcolor: red
urlcolor: blue
---

```{r, include = FALSE, echo = FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(ggplot2)
read_chunk('jasonCh2.R')
```

\newcommand{\rc}{\tikz\draw[black,fill=red] (0,0) circle (.5ex); }

\newcommand{\bc}{\tikz\draw[black,fill=blue] (0,0) circle (.5ex); }


# **Chapter 1: The Golem of Prague**

* Limitations of *deductive falsification*
  1. "Hypotheses are not models. The relations among hypotheses and different kinds of models are complex. Many models correspond
  to the same hypothesis, and many hypotheses correspond to a single model. This makes strict falsification impossible." (p. 4)
  2. "Measurement matters. Even when we think the data falsify a model, another ob- server will debate our methods and measures.
  They don’t trust the data. Sometimes they are right." (p.4)
* Main topics covered in this book...
  1. Bayesian data analysis
  2. Multilevel models
  3. Model comparison using information criteria

# **Chapter 2: Small Worlds and Large Worlds**

## **Useful Example.**

Suppose there is a bag containing marbles.  We *know* there are four marbles in the bag and that each marble
may be either \textcolor{red}{red} or \textcolor{blue}{blue}.

* *Prior belief:* The Bayesian approach always begins with a prior belief about the conjecture -- i.e., how many \textcolor{red}{red}
marbles are in the bag.  To incorporate a prior belief we must assign a plausiblity to each possibility.  Imagine that we met the owner
of the bag and they told us that there is a company that makes these bags and the do so in a way that ensure that there is a 15% chance
that a bag has only 1 \textcolor{blue}{blue} marble, a 50% chance that a bag as 2 \textcolor{blue}{blue} marbles, a 15% chance that a
bag has 3 \textcolor{blue}{blue} marbles, a 10% chance that a bag has all \textcolor{blue}{blue} marbles, and a 10% chance that the bag
does not contain any \textcolor{blue}{blue} marbles.  We could use this as our prior belief about the contents of our bag.

* *Data:*  Suppose we repeat the following steps 3 times: shake the bag (randomly distributing the marbles), blindly draw out a marble,
note the color, and put the marble back in the bag.  Now, suppose this exercise produced the following sequence of marbles \bc \rc \bc
in that exact order (*so sequence matters*).

* *Likelihood:* We can think of the likelihood as a way to count all of the possible ways of producing this sample **given** a
conjecture of what the bag looks like.Consider one possible conjecture: \bc \rc \rc \rc .  How likely is the observed sample
\bc \rc \bc given the conjecture?  To answer this question, we count the possible ways our conjecture can generate the sample.
Then we will compare the total number of ways across different conjectures as a way to evaluate which conjecture is the most likely
candidate for generating \bc \rc \bc .  The following plot shows all of the possible samples that (the conjecture) \bc \rc \rc \rc
can produce:

```{r, gardenPlot, echo = FALSE, out.width = '100%', fig.height=4, fig.width=6}
d
```
From this figure we see that there are 3 possible ways to generate the observed data. The following table lists all of the ways that
the 5 possible conjectures could generate the sample \bc \rc \bc:

| **Conjecture**   | **# of ways to produce** \bc \rc \bc |     **% of Total**    |
|:-----------------|:------------------------------------:|:---------------------:|
| \bc \rc \rc \rc | $1 \times 3 \times 1 = 3$             | $\frac{3}{20} = .15$  |
| \bc \bc \rc \rc | $2 \times 2 \times 2 = 8$             | $\frac{8}{20} = .40$  |
| \bc \bc \bc \rc | $3 \times 1 \times 3 = 9$             | $\frac{9}{20} = .45$  |
| \bc \bc \bc \bc | $4 \times 0 \times 4 = 0$             | $\frac{0}{20} = .00$  |
| \rc \rc \rc \rc | $0 \times 4 \times 0 = 0$             | $\frac{0}{20} = .00$  |

* *Bayesian updating & posterior distribution:* We update our prior information using the data, our likelihood, and multiplication.
The result is a probability distribution for our model parameters, which we can use to make inferences. \newpage

| **Conjecture**   | **Prior** | **Likelihood of** \bc \rc \bc | $\propto$ **Posterior**    | **Rescaled Posterior**     |
|:-----------------|:---------:|:-----------------------------:|:--------------------------:|---------------------------:|
| \bc \rc \rc \rc  | $.15$     |  $\frac{3}{20} = .15$         | $.15 \times .15 = .0225$   | $\frac{.0225}{.275} = .082$|
| \bc \bc \rc \rc  | $.50$     |  $\frac{8}{20} = .40$         | $.50 \times .40 = .2000$   | $\frac{.2}{.275} = .723$   |
| \bc \bc \bc \rc  | $.15$     |  $\frac{9}{20} = .45$         | $.15 \times .35 = .0525$   | $\frac{.0525}{.275} = .191$|
| \bc \bc \bc \bc  | $.10$     |  $\frac{0}{20} = .00$         | $.10 \times .00 = .0000$   | $\frac{0}{.275} = .000$    |
| \rc \rc \rc \rc  | $.10$     |  $\frac{0}{20} = .00$         | $.10 \times .00 = .0000$   | $\frac{0}{.275} = .000$    |

## **Another example of Bayesian updating.**

Consider a small globe that we toss up in the air (so that it spins) and when we catch
it we note if our right index finger is covering a portion of water (as opposed to land).  This is our parameter of interest,
i.e., the percent of the globe that is covered in water.  If our prior belief is that all of the probabilities are equally likely,
then the following figure shows what happens as we toss the ball, observe on outcome and update our prior beliefs.

Assume this is the sequence...W L W W W L W L W

```{r, globePlot, echo = FALSE, out.width = '100%', fig.height=4, fig.width=6}
g
```

## Evaluating posterior
grid, quadratic approximation, and MCMC.

# **Chapter 3**

# **Chapter 4**
